\LARGE \bf Artifical Cognition for Social Human-Robot Interaction:\\ An Implementation

### 2016-06-17 paper.tex
James' proof-reading of pages 13-14

\usepackage{xspace}
\newcommand{\ie}{{i.e.\xspace}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{e.g.\xspace}}
\newcommand{\etal}{{et al.\xspace}}

@@

    \caption{List of statements describing agent-independent spatial
        relationships between objects (top), agent-dependent placements (middle)
        and attentional states and abilities of agents (bottom).
        ``$\rightarrow$'' indicates sub-properties. When existing, the
        equivalent predicate in the {\sc OpenCyc} standard (prefix \concept{cyc:})
        is specified. Note that some relationships are not computed by {\sc Spark}, but
        are instead inferred by the reasoner.}

@@

Besides, temporal reasoning (essential for accurate action recognition for
instance) is not properly addressed in our architecture. Temporal reasoning is used
only locally, and does not allow for tracking of long sequences or global events.

@@

Natural language is a basic interaction modality that we use in our system both
as input (processing of the human speech) and as output (verbalization of the
robot intents). Natural language processing is facilitated by our architecture
mostly processing information with semantics that are already close to the
human-level. This section presents the main features of our speech processor,
{\sc Dialogs}, that include semantic and multi-modal grounding, and interactive
disambiguation. Algorithmic and implementation details are provided
in~\cite{Lemaignan2011a}.

We acquire natural speech input from the human participants through a custom
Android-based interface that relies on the Google speech recognition API for
speech-to-text and relays the textual transcript to the robot. It is parsed into
a grammatical structure (\emph{Part of Speech} tagging), and the atoms of each
sentence are resolved with the help of the ontology to ground concepts like
objects (\ie when a user says ``pick the can'', it resolves to which instance of
\emph{can} the user is referring to) and actions.  Figure~\ref{dialogs|ex} gives
an example of the processing of a simple, non-ambiguous command, while the first
study (Section~\ref{moving-london}) walks through more complex examples.
Heuristics (like the presence of a question mark or the use of imperative mood)
are used to classify the sentences into questions, desires or statements, and
{\sc Dialogs} processes them accordingly (by answering questions or
updating the knowledge base).

@@

    \caption{Processing of a simple, non-ambiguous command, taken
        from~\cite{Lemaignan2011a}. Thematic roles ({\tt performedBy}, {\tt
        actsOnObject}, {\tt receivedBy}) are automatically extracted by 
    {\sc Dialogs} from the imperative sentence ``Give me the book.'' The new
    resulting statements (right column) are added to the knowledge base, and
    may eventually trigger an event in the execution controller (see
    Section~\ref{events}).}

@@

The system supports quantification (``give me \{a | the | some | all | any |
...\} can''), thematic roles (action-specific predicates that qualify the
actions), interactive disambiguation (the robot asks questions when it needs
more information), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also supports knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal} and added to the knowledge base
after checking for possible contradictions with existing knowledge),
interprets common temporal and spatial adverbs (like \emph{above} or
\emph{tomorrow}) and translates simple expressions of internal state (``I'm
tired'') into \emph{experiences} (\stmt{human\_1 experiences state\_1, state\_1
hasFeature tired}, see also Section~\ref{sect:desires}). A full account of the
{\sc Dialogs} features and the corresponding algorithms is available
in~\cite{Lemaignan2011a}.

@@

Because the components of our architecture rely on the same RDF formalism to
represent their outputs, the different communication modalities (\emph{explicit}
like verbal, deictic or based on gaze, or \emph{implicit} like the position of
the human) are presented in a homogeneous way, as statements in the knowledge
base. The dialogue grounding process makes use of them at two distinct levels to
provide multi-modal concept grounding.

First, specific steps of the grounding process explicitly check for the
presence and value of specific facts: for instance, when several instances match
a category (the human says ``give me the bottle'' and the robot knows about
three bottles), the module may decide (it actually depends on the quantifier
preceding the class, see the example at section~\ref{moving-london}) to discard
some of them based on their \emph{visibility} for the speaker (implicit
communication context taking into account the human position).

Another example, when the human says ``this'', the robot checks if the human is
currently pointing at some object. In that case, \emph{this} is replaced by the
object focused on. Otherwise, the robot performs a lookup in the dialogue history
to find a previous concept that the user could refer to (anaphora resolution).

Note that, while the system benefits from the complementary modalities, they are
not all required. The dialogue system can run with the verbal modality alone, at
the cost of a simpler interaction: if the human says ``this'' without the robot
tracking what the human points at, no \stmt{human\_1 pointsAt ...} fact would
possibly be available in the knowledge base, and the robot falls back to the
anaphora resolution step alone.

The second level of integration of multi-modality is implicit. By continuously
computing symbolic properties from the geometric model, richer descriptions and
hence discrimination possibilities are available. For instance, the robot may
compute that one bottle is next to a glass while another one stands alone, and
these symbolic descriptions are transparently re-used in a dialogue context to
generate unambiguous references to discriminate between similar objects: ``do
you mean the bottle that is next to the glass?''. \cite{Ros2010b} provides a
detailed account of our approach to interactive concept clarification and
discrimination, along with the related algorithms.

@@

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. We use in our architecture the HATP
planner (\emph{Human Aware Task Planner})~\cite{Alili2008,
Alili2009,Lallement2014}.

The HATP planning framework extends the traditional HTN (Hierarchical Task
Network) planning domain representation and semantics by making them more
suitable to elaborate plans which involve humans and robots acting together
toward a joint goal. HATP is used by the robot to elaborate human-robot
\emph{shared plans}~\cite{Grosz1996,Clark1996,Kemp2007} which are then used to
anticipate human action, to suggest a course of action to humans, or even to ask
help from the human.

@@

%\paragraph{Action Execution and Monitoring Robot Controller}
%Based on context and on the shared plan elaborated by HATP for a given goal,
%the robot controller decides to execute an action or to ask its human
%partner to do it.  Actions feasibility by the human or the robot are
%regularly reconsidered based on the reachability / visibility
%computation mechanisms.
%
%Robot action execution is based on simple automatons  that translate
%symbolic planning atomic actions into sequences of planned arm motions
%and gripper commands to execute according to current state
%of the 3-tuple (gripper, object, furniture). We have three states
%according to whether the object is in gripper and if it is in gripper
%whether it is on furniture.  These states are directly obtained from
%the updated symbolic state of the world in the ontology.
%
%For plan action monitoring, primitive actions recognition is used. A primitive
%action detection is interpreted as action success if it is the expected one and
%failure otherwise. The robot also reacts to the absence of activity.
%

@@

Motion and action execution also requires human models, and the one we use
embeds human preferences and physical constraints that need to be accounted for
when synthesizing robot motion or elaborating robot plans. This includes
proxemics (human-robot distance) and associated issues (visibility) but also
legibility and acceptability criteria expressed in terms of \emph{social rules}
that the elaborated plans should satisfy.

>>>>

\usepackage{xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{e.g.\xspace}
\newcommand{\etal}{et al.\xspace}

@@

    \caption{List of statements describing agent-independent spatial
        relationships between objects (top), agent-dependent placements
        (middle), and attentional states and abilities of agents (bottom).
        ``$\rightarrow$'' indicates sub-properties. Where existing, the
        equivalent predicate in the {\sc OpenCyc} standard (prefix
        \concept{cyc:}) is specified. Note that some relationships are not
        computed by {\sc Spark}, but are instead inferred by the reasoner.}

@@

Additionally, temporal reasoning (essential for accurate action recognition for
instance) is not properly addressed in our architecture. Temporal reasoning is used
only locally, and does not allow for tracking of long sequences or global events.

@@

Natural language is a basic interaction modality that we use in our system both
as an input (processing of the human speech) and as an output (verbalization of the
robot intentions). Natural language processing is facilitated by our architecture
that processes information with semantics that are already close to a
human-level. This section presents the main features of our speech processor,
{\sc Dialogs}, that include semantic and multi-modal grounding, and interactive
disambiguation. Algorithmic and implementation details are provided
in~\cite{Lemaignan2011a}.

We acquire natural speech input from the human participants through a custom
Android-based interface. The interface relies on the Google speech recognition API for
speech-to-text and relays the textual transcript to the robot. The text is parsed into
a grammatical structure (\emph{Part of Speech} tagging), and the atoms of each
sentence are resolved with the help of the ontology to ground concepts like
objects (\ie when a user says ``pick up the can'', it resolves to which instance of
\emph{can} the user is referring to) and actions.  Figure~\ref{dialogs|ex} gives
an example of the processing of a simple, non-ambiguous command. The first
study (Section~\ref{moving-london}) walks through more complex examples.
Heuristics, like the presence of a question mark or the use of imperative mood,
are used to classify the sentences into questions, desires or statements.
{\sc Dialogs} processes these accordingly by answering questions or
updating the knowledge base.

@@

    \caption{Processing of a simple, non-ambiguous command, taken
        from~\cite{Lemaignan2011a}. Thematic roles ({\tt performedBy}, {\tt
        actsOnObject}, {\tt receivedBy}) are automatically extracted by 
    {\sc Dialogs} from the imperative sentence ``Give me the book.'' The 
    resulting statements (right column) are added to the knowledge base, and
    may eventually trigger an event in the execution controller (see
    Section~\ref{events}).}

@@

The system supports quantification (``give me \{a | the | some | all | any |
$n$\} can''), thematic roles (action-specific predicates that qualify the
actions), interactive disambiguation (the robot asks questions when it needs
more information), and anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also supports knowledge extension by learning new semantic
structures. For instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal} and added to the knowledge base
after checking for possible contradictions with existing knowledge. {\sc
Dialogs} finally interprets common temporal and spatial adverbs (like
\emph{above} or \emph{tomorrow}) and translates simple expressions of internal
state  into \emph{experiences} (for instance, ``I'm tired'' is processed into
\stmt{human\_1 experiences state\_1, state\_1 hasFeature tired}, see also
Section~\ref{sect:desires}). A full account of the {\sc Dialogs} features and
the corresponding algorithms is available in~\cite{Lemaignan2011a}.

@@

Because the components of our architecture rely on the same RDF formalism to
represent their outputs, the different communication modalities are presented in
a homogeneous way, as symbolic statements in the knowledge base. This applies
both to \emph{explicit} modalities (verbal communication, deictic gestures,
social gaze), and \emph{implicit} modalities (like the body position of the
human). The dialogue grounding process makes use of them at two distinct levels
to provide multi-modal concept grounding.

First, specific steps of the grounding process explicitly check for the presence
and value of certain facts. For instance, when several instances match a
category (the human says ``give me the bottle'' and the robot knows about three
bottles), the module may decide to discard some of the candidates based on their
\emph{visibility} for the speaker (implicit communication context taking into
account the human position). In this particular case, the heuristic is selected
by {\sc Dialogs} based on the quantifier preceding the class (``give me
\underline{the} bottle''). The first study (Section~\ref{moving-london})
illustrates the details of this process.

As another example, when the human says ``this'', the robot checks if the human
is currently pointing at an object. In that case, \emph{this} is replaced by the
object focused on. Otherwise, the robot performs anaphora resolution by looking
up in the dialogue history to find a previous concept that the user could refer
to.

Note that while the system benefits from complementary modalities, they are not
all required. The system can run with the verbal modality alone, at the cost of
a simpler interaction. For example, if the human says ``this'' without the robot
tracking what the human points at, no \stmt{human\_1 pointsAt ...} fact would
possibly be available in the knowledge base, and the robot falls back on the
anaphora resolution step alone.

The second level of integration of multi-modality is implicit. By continuously
computing symbolic properties from the geometric model, richer symbolic
descriptions are available to the system to verbalize or discriminate entities.
For instance, the robot may compute that one bottle is next to a glass, while
another one stands alone. These symbolic descriptions are transparently
re-used in a dialogue context to generate unambiguous references to discriminate
between similar objects: ``do you mean the bottle that is next to the glass?''.
The physical context of the interaction is used as an implicit communication
modality by {\sc Dialogs}. \cite{Ros2010b} provides a detailed account of our
approach toward interactive concept clarification and discrimination, along with the
related algorithms.

@@

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. Our architecture uses the HATP
planner (\emph{Human Aware Task Planner})~\cite{Alili2008,
Alili2009,Lallement2014}.

The HATP planning framework extends the traditional Hierarchical Task
Network (HTN) planning domain representation and semantics by making them more
suitable to produce plans which involve humans and robots acting together
toward a joint goal. HATP is used by the robot to produce human-robot
\emph{shared plans}~\cite{Grosz1996,Clark1996,Kemp2007} which are then used to
anticipate human action, to suggest a course of action to humans, or even to ask
help from the human.

@@

%\paragraph{Action Execution and Monitoring Robot Controller}
%Based on context and on the shared plan produced by HATP for a given goal,
%the robot controller decides to execute an action or to ask its human
%partner to do it.  Actions feasibility by the human or the robot are
%regularly reconsidered based on the reachability / visibility
%computation mechanisms.
%
%Robot action execution is based on simple automatons  that translate
%symbolic planning atomic actions into sequences of planned arm motions
%and gripper commands to execute according to current state
%of the 3-tuple (gripper, object, furniture). We have three states
%according to whether the object is in gripper and if it is in gripper
%whether it is on furniture.  These states are directly obtained from
%the updated symbolic state of the world in the ontology.
%
%For plan action monitoring, primitive actions recognition is used. A primitive
%action detection is interpreted as action success if it is the expected one and
%failure otherwise. The robot also reacts to the absence of activity.
%

@@

Motion and action execution also requires human models, and the one we use
embeds human preferences and physical constraints that need to be accounted for
when synthesizing robot motion or producing robot plans. This includes
proxemics (human-robot distance) and associated issues (visibility) but also
legibility and acceptability criteria expressed in terms of \emph{social rules}
that the produced plans should satisfy.
